---
title: "P8105_hw3_wl2829"
author: "Wentong"
date: "10/15/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

Load the data.
```{r}
library(p8105.datasets)
data("instacart")

```
The instacart dataset contains 1384617 rows and 15 variables. This data comes from shopping websites, showing the sales of different products in different regions. The detail of information of product is clearly showed by variables such as aisle, product_name and product_id. We can also know the user and when the order was set. For example, the Bulgarian Yogurt in dairy eggs department was bought by user 112108. This order was nine days after the previous order and was finished on the 4th day of the week. 


### How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>% 
  count(aisle_id) %>% 
  nrow()   ## The total number of aisles.

instacart %>% 
  distinct() %>% 
  count(aisle, name = "aisle_n") %>% 
  mutate(aisle_n = as.numeric(aisle_n)) %>% 
  arrange(desc(aisle_n)) %>% ## rank the number of aisles
  filter(aisle_n == max(aisle_n)) ## pick the biggest number
```
"Fresh vegetables" has the most items. The number of orders is 150609.

### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
plot_df = 
  instacart %>% 
  distinct() %>% 
  count(aisle, name = "aisle_n") %>% 
  mutate(aisle_n = as.numeric(aisle_n)) %>% 
  filter(aisle_n >= 10000) %>% ## remove the data less than 10000
  arrange(desc(aisle_n)) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = forcats::fct_reorder(aisle, sort(aisle_n))) %>% 
  ## control the factor
  ggplot(aes(x = aisle, y = aisle_n)) + 
  geom_point() + 
  coord_flip() +
  labs(
    title = "Number of items more than 10000",
    y = "number of itmes",
    x = "name of items",
    caption = "Data from instacart; number of items."
  ) ## make the plot readable

plot_df
```
We can see from the plot that fresh vegetables and fresh fruits are the most popular department in instacart. The number of most product is less than 40000.

### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
Bi_df = 
  instacart %>% 
  filter(aisle == "baking ingredients") %>%   ## select the specific aisle
  count(product_name, name = "product_n") %>% ## count the number of each product
  arrange(desc(product_n)) %>% 
  head(3) %>% ## order the number and pick the three most popular items
  mutate(aisles_name = "baking ingredients") %>% 
  select(aisles_name, product_name, product_n) ## rearrange the table

dfc_df = 
  instacart %>% 
  filter(aisle == "dog food care") %>% 
  count(product_name, name = "product_n") %>% 
  arrange(desc(product_n)) %>% 
  head(3) %>% 
  mutate(aisles_name = "dog food care") %>% 
  select(aisles_name, product_name, product_n)

pvf_df = 
  instacart %>% 
  filter(aisle == "packaged vegetables fruits") %>% 
  count(product_name, name = "product_n") %>% 
  arrange(desc(product_n)) %>% 
  head(3) %>% 
  mutate(aisles_name = "packaged vegetables fruits") %>% 
  select(aisles_name, product_name, product_n)

three_aisles_df = 
  rbind(Bi_df, dfc_df, pvf_df) %>% 
  knitr::kable()

three_aisles_df
```

### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
mean_hour_df = 
  instacart %>% 
  filter(product_name %in% c("Coffee Ice Cream", "Pink Lady Apples")) %>% 
  ## select the data
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_hour = mean(order_hour_of_day) 
  ) %>% ## calculate the mean of the data
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% ## format the data to a 2 x 7 table
  rename("Sun" = "0","Mon" = "1", "Tue" = "2", "Wed" = "3", "Thu" = "4", 
         "Fri" = "5", "Sat" = "6") %>%  ## convert number to weekday name
  knitr::kable()

mean_hour_df
```


## Problem 2
Load the data
```{r}
library(p8105.datasets)
```

```{r}
BRFSS_df = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr, specific_location = locationdesc) %>% 
  ## format the data to use appropriate variable names
  filter(topic == "Overall Health") %>% ## focus on the “Overall Health” topic
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  ## include only responses from “Excellent” to “Poor”
  mutate(response = factor(response), 
         response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent"))) 
  ## organize responses as a factor taking levels ordered from “Poor” to “Excellent”

BRFSS_df
```

### In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
BRFSS_df %>% 
  filter(year == 2002) %>% 
  select(state, specific_location) %>% 
  distinct() %>% 
  count(state, name = "ob_n") %>% 
  filter(ob_n >= 7) %>% 
  knitr::kable()
## states were observed more than 7 in 2002


BRFSS_df %>% 
  filter(year == 2010) %>% 
  select(state, specific_location) %>% 
  distinct() %>% 
  count(state, name = "ob_n") %>% 
  filter(ob_n >= 7) %>% 
  knitr::kable()
## states were observed more than 7 in 2010
```

### Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state.

```{r}
BRFSS_df %>% 
  filter(response == "Excellent") %>% 
  group_by(year, state) %>% 
  mutate(
    mean_value = mean(data_value, na.rm = TRUE)
  ) %>% ## calculate the average value
  select(year, state, mean_value) %>% 
  distinct() %>% 
  ggplot(aes(x = year, y = mean_value)) +
  geom_line(aes(group = state, color = state))

```

### Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses among locations in NY State.

```{r}
two_panel_plot = 
  BRFSS_df %>% 
  filter(year %in% c(2006, 2010)) %>% 
  ## select the 2006 and 2010 data
  filter(state == "NY") %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ year)
  ## make a two-panel plot

two_panel_plot
```

## Problem 3

### Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes.
```{r}
acce_df = 
  read_csv("./data/accel_data.csv") %>% ## load the data
  pivot_longer(
    activity.1:activity.1440,
    names_to = "minute_of_activity",
    values_to = "activity"  
  ) %>% ## reduce the number of variables to make the data more clear
  janitor::clean_names() %>% 
  mutate_at("minute_of_activity", str_replace, "activity.", "") %>%
  ## remove the useless part of the data
  mutate(day_type = case_when(
    day %in% c("Friday", "Monday", "Tuesday", "Wednesday", "Thursday") ~ "weekday",
    day %in% c("Saturday", "Sunday") ~ "weekend"
  )) %>%  ## add a weekday vs weekend variable
  mutate(minute_of_activity = as.numeric(minute_of_activity))
  ## encode data with reasonable variable classes

acce_df
```

### Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
acce_df %>% 
  group_by(day_id) %>% 
  mutate(day_sum = sum(activity)) %>% 
  ## create a total activity variable for each day
  select(day_id, day_type, day, day_sum) %>% 
  distinct() %>% 
  knitr::kable()
```
he data in the table does not show a clear trend of data changes, but the total amount in certain days is relatively low, especially on Saturdays in the fourth and fifth weeks

### Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. 


```{r}
acce_df %>% 
  mutate(minute_of_activity = as.numeric(minute_of_activity)) %>% 
  ggplot(aes(x = minute_of_activity/60, y = activity, color = day)) +
  geom_smooth() 
```

This plot shows that the fluctuation of each day's data is very obvious. Activities value are relatively low between 12:00pm to 6:00am. However, after 6:00am, the value of the data began to rise and reaches the maximum at 9:00pm. Among them, the data of Saturday is quite special, since it reaches the maximum at about 10:00am. 
